\chapter{Problem justification}

The use of social media has resulted in profound changes in the pattern of human interaction in today's contemporary society. The presence of social media has emerged as a phenomenon that places itself as a crucial element in the drastic changes in the way people interact in an era where information technology has undeniable dominance in almost every field of life \parencite{azzaakiyyah2023impact}. As these platform become central to public discourse and information dissemination, the methods used to analyze them face increasing scrutiny. However, much of the existing analysis is conducted using proprietary, \enquote{black-box} tools that limit transparency and prevent academic certification. This creates a critical need for \textbf{an application built on a reproducible, auditable, and documented methodology} for a major platform like YouTube, thereby empowering researches and the public with verifiable insights.

\section{Academic}

This project directly addresses a critical methodological gap in academic research: the lack of a unified, verifiable system for YouTube data analysis. Its core academic contribution is the promotion of \textbf{reproducibility} and \textbf{methodological transparency}. By providing a \textbf{fully documented methodology within this thesis}, the application's architecture and data pipeline allow for the independent verification of research findings. This application provides a \textbf{foundational tool}, enabling researchers to conduct reproducible analyses and serving as a \textbf{validated base for future work} in more advanced data analysis. This creates a valuable and reusable asset for the academic community, ensuring that future topic-level studies of YouTube can be conducted with higher methodological rigor and transparency.

Concretely, this application will support active research initiatives such as the PROMUEVA project at Universidad del Valle, which requires social media data to develop computational models for analyzing polarization in Valle del Cauca. By providing a transparent data collection infrastructure, this work enables such projects to access YouTube data with full methodological documentation, supporting their research on polarization patterns and user behaviors in social networks.

\section{Economic}

The primary economic benefit is providing a \textbf{no-cost, high-value alternative} to expensive social listening tools. Commercial tools like Talkwalker or Meltwater represent a significant financial barrier for academic institutions, non-profits and independent researchers, limiting their ability to conduct large-scale analysis. By developing a \textbf{free tool}, this project \textbf{democratizes access} to these analytical capabilities, providing functionality that is otherwise locked behind costly subscriptions. Furthermore, the application \textbf{automates the entire data pipeline}, from collection and cleaning to analysis and visualization. This automation creates significant efficiency gains, saving countless hours of manual labor that researchers would otherwise spend on data preparation.

\section{Social}

Given that a substantial portion of the population now uses YouTube as a primary source for news and information, understanding the discourse on this platform is a matter of public interest. This project contributes to social good offering a \textbf{transparent} tool for analyzing public opinion, which can be used by academic researchers. Unlike proprietary \enquote{black-box} systems, this tool's \textbf{publicly documented methodology} ensures that its processes are completely \textbf{transparent} and \textbf{auditable}. This fosters greater trust and empowers organizations to monitor topics like misinformation, public health discussions or political sentiment with a \textbf{verifiable} and \textbf{accessible} tool.
