\section{Project Scope}

\subsection{Inclusions}

This section details the specific features, deliverables, and technologies that are part of the project.

\subsubsection{Functionality}

The project will deliver a fully functional web application that automates the collection, processing, analysis, and visualization of data from YouTube. The core functionalities include:

\begin{itemize}
	\item \textbf{Data collection module}: a configurable module will be implemented to collect video and associated comment data directly from the YouTube Data API using specified query filters.

	\item \textbf{AI query module}: a natural language interface that integrates with a pre-trained AI model to translate informal user requests into valid YouTube Data API queries.

	\item \textbf{Data processing pipeline}: the application will feature a processing pipeline designed to clean, normalize, validate, and structure the raw data retrieved from the API.

	\item \textbf{Presentation layer}: a user-facing presentation layer will be built, featuring interactive dashboards and charts to visualize the analyzed data.
\end{itemize}

\subsubsection{Technology}

\textbf{The specific technology stack for the application will be selected and justified as part of the implementation phase}. However, the architecture will adhere to the following principles:

\begin{itemize}
	\item \textbf{Backend service}: the core application logic and API will be developed using a modern web framework, with a relational database for data storage.

	\item \textbf{Frontend service}: a lightweight web server (e.g., Nginx) will be used to serve the static frontend assets (HTML, CSS, and JavaScript) built from a modern JavaScript framework (e.g., React, Vue).

	\item \textbf{Containerization}: all services (web server, backend application, database) will be containerized using Docker to ensure a consistent and reproducible deployment environment.
\end{itemize}

\subsection{Exclusions}

To maintain a clear focus and ensure the project is achievable within the defined timeframe, the following features and functionalities are explicitly out of scope:

\begin{itemize}
	\item \textbf{Support for other platforms}: data collection and analysis will be limited exclusively to YouTube. The application will not support any other social media or video platforms (e.g., Facebook, Twitter, TikTok).

	\item \textbf{Training of NLP models}: the project will \textbf{integrate} existing, pre-trained models (for sentiment analysis and query translation). The scope does not include the \textbf{training or fine-tuning} of new or custom natural language processing or machine learning models from scratch.

	\item \textbf{Real-time data streaming}: the data collection process is designed to be on-demand based on user queries. The project will not implement real-time or live-streaming data analysis.

	\item \textbf{Native mobile application}: the final deliverable is a web application accessible through standard desktop browsers. The development of native mobile applications for iOS or Android is not included.
\end{itemize}
